% Using \unnumberedsection here for now (instead of \appdxsection and
% then \appdxsub... within) because appendix numbering seems to be
% broken when we use \unnumberedsection for all the previous sections.
\unnumberedsection{Appendix C: Ecosystem Mapping Worksheet}\label{appdx-worksheet-ecosystem-map}

Use this page to draw a map or directed graph of current and potential
actors in your ecosystem.  List service providers and group them by the
type of service they offer.  Identify potential collaborators, and
mark the ones with competitive service offerings.  Identify competing
open source and proprietary substitutes for your open source project.

\vspace{1em}

Place actors with large, current impact closer to the center of the
map and future recruits further away.  The open source project belongs
at the center; your company might be close in or further out,
depending on its current effective scale of investment.

\vspace{1em}

Don't aim for a perfect representation of the project's ecosystem at
first.  Just get lots of things down on paper --- then you can
rearrange, regroup, recategorize, and edit them as needed.  When done,
note interesting relationships between various nodes on the map.  Add
customers in another color.  This map is a picture of your world as it
currently exists and how it might change in the near-term future.  Be
sure to save a snapshot of this map and see how it shifts over time.

% It would be good to have an example here.
% 
% https://twitter.com/dlshadothman/status/1065317596440989699
% is an example of an ecosystem map, but I'm not sure it's
% the right example for our purposes:
% 
% 1) It's based on publicly available data, whereas a real ecosystem
%    map would probably include privately-known information or
%    information that one is willing to guess about in private that
%    one wouldn't be as comfortable guessing about in public.  
%
% 2) It doesn't name very many names.  There are a few, but mostly it
%    talks in the abstract about communities, governments, etc.  That
%    can be useful for archetypes analysis too, but I think we want an
%    example that demonstrates the naming of specific actors.
%
% 3) It's done as a large, complex Venn diagram, which is a bit
%    limited in comparison to the ecosystem maps we usually use.
%
% James said of this map:
%
%   "It includes mostly the project and a little bit of its
%   connections with the outside world, but does not include all the
%   other projects that are related to Tor.  However, an ecosystem map
%   that was more widely scoped would be useful for other purposes.
%   With more complex projects, the need for a project-specific
%   ecosystem map increases."
