\subsection{Specialty Library}\label{archetype:specialty-library}

{\bf Examples:} \emph{libssl, libmp4}

{\bf Characteristics:} Specialty libraries provide base functionality to a
set of functions whose implementation requires specialized knowledge
at the leading edge of research.  E.g., developers shouldn't roll
their own crypto or codecs; instead, everyone benefits from easy
access to a pooled set of code built by experts.

These libraries are fairly commoditized, as the set of capabilities for
such a library is fairly fixed.  Any ssl or mp4 library is going to do
roughly the same thing.  Competing libraries might differ at the API
level, but there is no arms race over features.  Given the lack of
opportunity for competitive advantage in developing one's own library,
most of those capable of contributing have every reason to do it as
part of an open source project that shares pooled development costs
and benefits.

Widespread adoption of a specific open source library can power
standardization work and also increase adoption of formal and informal
standards.  Coupled with careful approaches to patents, this can open
up access to participants who aren't attached to companies with deep
pockets and large patent portfolios.

Specialty libraries look like open source projects with high barriers
to entry.  Anybody can submit a patch, but landing it might require a
high degree of expert work before it is accepted into core.  Coding
standards are high.  There is
often less developer outreach and hand-holding of new developers as
participants are largely assumed to be experienced and capable.  In a
heavily patented area or one subject to export controls, acceptance
policies might be more complex and even involve legal review.

\begin{itemize}
\item {\bf Licensing}: Usually non-copyleft.
\item {\bf Community standards}: High barriers to entry, largely
  because contributors are expected to be experts.
\item {\bf Component coupling}: Tightly coupled.  These libraries are
  structured to do one thing well.
\item {\bf Main benefits}: Ensures a shared solution to a specific
  technical problem.  Cooperation at the engineering level can lead to
  new organizational partnerships.
\item {\bf Typical governance}: Formal committer group that can
  grant committing privileges to new contributors.
\end{itemize}
