\subsection{Business-to-Business (B2B) Open Source}\label{archetype:b2b-open-source}

{\bf Examples:} \emph{Android, Chromium}

{\bf Characteristics:} Solid control by founder
and (sometimes grudging) acceptance of unpredictability by
redistributors.  This also implies a long fork-modify-merge cycle for
partners, and thus a constant temptation on their part to fork
permanently.

This archetype aims at ubiquity.  Its open source nature is designed
to spur OEM adoption by partners and competitors across the industry
and to drive out competing options in the same way pricing browsers at
zero once discouraged new entrants into the browser market.  It can
also drive down the price of complementary products: for example, as
Red Hat Enterprise Linux (RHEL) does for server hardware and Google's
Android operating system does for handheld devices.\footnote{Compare
  with the Mass Market archetype, which can have similar effects on
  directly competitive products but is less likely to affect
  complementary products.}

It is worth noting that Google does not derive much direct revenue
from Android.  Instead, Android's popularity puts Google's
products and search preferences in the hands of users, and that
creates opportunities for Google.  Android users default to Google's
search engine, buy media from Google, pay for apps in Google's app
store, provide a river of data for Google to mine, and favor Google's
app ecosystem (Calendar, Gmail, Maps, etc).  All of that generates
revenue and strategic advantage for Google.  This archetype is thus a
strategy for gaining marketshare as a \emph{revenue opportunity}.

This model probably works best when run by a fairly large company with
multiple ways of applying market pressure and multiple channels for
distribution.  It is difficult to wield without considerable financial
resources and other strategic advantages.  \otsref{(ref:c25750f5)}

\begin{itemize}
\item {\bf Licensing}: Almost always non-copyleft.
\item {\bf Community standards}: In general, the lead company does not
  put much emphasis on welcoming or nurturing contributors; exceptions
  may be made for strategically important organizational partners.
\item {\bf Component coupling}: Tightly coupled modules, to allow the
  lead company to market one unified product.
\item {\bf Main benefits}: Can drive industry adoption of a technology
  that is strategically important to your organization.
\item {\bf Typical governance}: Maintainer-led by a group within
  the lead company.
\end{itemize}
