\hypertarget{archetype:mass-market}{}
\subsection{Mass Market}\label{archetype:mass-market}

{\bf Examples:} \emph{Firefox, LibreOffice, MediaWiki (due to Wikipedia instance)}

{\bf Characteristics:} Mass Market projects are often similar to
``Wide Open'' projects, but they necessarily have a protective layer
around their contribution intake process.  This is because there are
simply too many users and too many opinions for everyone to be heard
by the development team, and because the average technical
sophistication of the users is low (especially in terms of their
familiarity with the conventions of open source participation).  This is
especially true in a commercial context where product experience and
time-to-market are highly sensitive.

Like B2B, Mass Market tends to drive down the price of directly
competitive products, the way OpenOffice.org (and later LibreOffice)
did with Microsoft Office.  But Mass Market is less likely than B2B to
affect complementary products, because Mass Market is usually aimed at
individuals not at intermediary organizations.

Mass Market open source projects have certain opportunities that come
with economies of scale, and they should actively seek to take
advantage of those opportunities.  For example, they can often get
good translation (i.e., localization) coverage, because there are many
users who are qualified and motivated to perform translation even
though they might not have the skills to participate as programmers.
Mass Market projects can do effective A/B testing and user surveys,
and have some ability to influence \foreignphrase{de facto} or formal
Internet and Web standards.  Finally, they can be thought leaders in a
broader political sense (e.g., through industry partnerships, user
clubs, etc).

Mass Market might at first seem more like a description of an end
product than of a project archetype, but in fact there are important
ways in which being Mass Market has implications for how the project
is run.  With developer-oriented software, it usually pays to listen
to the most vocal customers.  But Mass Market projects are oriented
mainly toward non-technical users, and thus should be guided more by
explicit user research.
% (ref:b7de5e0a)

\begin{itemize}
\item {\bf Licensing}: Non-copyleft generally, but may be copyleft
  depending on the project's business strategy.
\item {\bf Community standards}: Fully open, but relatively brusque
  for the vast majority of users.  Because of the large number of
  users, these projects evolve toward a users-helping-users pattern,
  rather than the development team serving as user support.  The team
  itself can often seem somewhat distant or unresponsive.
\item {\bf Component coupling}: Mix and match packages.  This
  archetype can support a number of different component packaging
  techniques.
\item {\bf Main benefits}: Large user base can help the project be
  broadly influential.
\item {\bf Typical governance}: Technical committee and/or formal
  committer group.
\end{itemize}
