\hypertarget{archetype:controlled-ecosystem}{}
\subsection{Controlled Ecosystem}\label{archetype:controlled-ecosystem}

{\bf Examples:} \emph{WordPress, Drupal}

{\bf Characteristics:} Real community involvement, and diversity of
community motivations, but with a shared understanding that the
founder (or some credible entity) will act as ``benevolent dictator''.
Requires experienced open source community management on part of
project leaders and occasional willingness to compromise.  Works best
when technical architecture directly supports out-of-core innovation
and customization, such as via a plugin system.

Controlled Ecosystem efforts find much of their value in that
ecosystem.  The core provides base value, but the varied contributions
across a healthy plugin ecosystem allow the project to address a much
larger and diverse set of needs than any one project could tackle
alone.  Over time, these projects might see more and more of the core
functionality structured as plugins as the product becomes more
customizable and pluggable.  This increasingly lets the plugins
determine the end-user experience, and the core project can eventually
become infrastructure.

\begin{itemize}
\item {\bf Licensing}: Can be either copyleft or non-copyleft.
  When copyleft, decisions must be made about whether the core
  interfaces (and maintainer-promulgated legal interpretations)
  encourage or discourage proprietary plugins.
\item {\bf Community standards}: Welcoming, often with structures
  designed to promote participation and introduce new contributors.
\item {\bf Component coupling}: Loosely coupled modules, frequently in
  a formal plugin system.
\item {\bf Main benefits}: Builds a sustainable ecosystem in which the
  founding organization retains strong influence.
\item {\bf Typical governance}: Benevolent dictatorship, with
  significant willingness to compromise to avoid forks.
\end{itemize}
