\hypertarget{business-models}{}
\unnumberedsection{Business Models and Archetypes}\label{business-models}

A project's archetype is just one component of some larger strategy,
and the strategy should drive the archetype, not the other way around.
To think ``Okay, we have a Rocket Ship To Mars project here.  What
business models can it support?'' would be putting the cart before the
horse.  The better way is to start from strategy: ``The business model
we see this project being part of is X.  What archetypes work well
with X?''  (There may be more than one business model, of course, just
as a project may end up incorporating elements from multiple
archetypes.)

Different archetypes suit different business models, similarly to how
different open source licenses suit different kinds of for-profit
activity.  In licensing, there are times when a
non-copyleft\footnote{A.k.a. ``permissive''.  See
  \fullref{licenses-variation}, p. \pageref{licenses-variation}.}
license is the right choice, given
your intended business model (perhaps widespread adoption is crucial
to your plans), and there are times when the strongest possible
copyleft\footnote{E.g., Currently the AGPL.  As of this writing in
  early 2019, the debate over whether the newly-created ``Server Side
  Public License'' from MongoDB is or is not a free software / open
  source license is still very new.  Because it has not yet been
  evaluated by the OSI nor declared a free software license by the
  FSF, we take a conservative approach and do not consider it here.
  This does not imply any position on whether it is or isn't a free
  and open source (FOSS) license.  It simply acknowledges the
  practical reality that SSPL-licensed software is not going to be
  widely treated as FOSS unless and until the relevant certifying
  organizations treat it as FOSS.  See
  \fullref{appdx-licensing-basics},
  p. \pageref{appdx-licensing-basics}, for more context.} is the right
choice (perhaps avoidance of proprietary derivatives is crucial to
your plans).

Similar considerations apply when choosing an archetype.  If gaining
mindshare among individual developers is crucial, then you need an
archetype (such as Wide Open) that prioritizes participant onboarding,
even if it means sacrificing development speed or other technical
goals.  If building a coalition to forestall market domination by a
large competitor is your goal, then an archetype designed for
inter-organizational collaboration (such as B2B or Multi-Vendor
Infrastructure) is a better choice.

Think of ``business model'' as a superset of ``revenue model''.  A
revenue model is just one part of a business model.  You must start
from a clear definition of your product and your value proposition.
For example, in proprietary software, businesses that look at first
glance like they are based on a royalty-per-copy revenue model often
turn out, on closer examination, to be more complex than that.
Sometimes a business \emph{thinks} it is selling per-seat or
per-server usage rights, while its customers think they're what
they're paying for is support, vetted installation, configuration
convenience, and regular security and feature updates.

Most revenue models are compatible with open source, in general.
Really, the only one that is incompatible with open source is the one
based on per-copy royalties for software sales.  While that can be a
very successful\footnote{If you define success primarily in terms of
  revenue generation, that is.  For mission-driven organizations,
  success involves more than that anyway.} revenue model under certain
circumstances, it is far from the only one.

This report is not the place for an in-depth discussion of value
creation, value capture, market definition, and revenue models.  We
note these complex topics merely to point out that any given open
source archetype will be better-suited to some business models and
worse-suited to others.  In order to determine which archetype is best
for your project, you must start from the high-level strategy --- in
which revenue is just one element --- that motivated your involvement
in project in the first place.
