\numberedsection{Contributors and Archetypes: Setting Expectations}\label{contributor-expectations}

Knowing a project's archetype helps to answer an important question:

\begin{adjustwidth}{3em}{3em}
\emph{How much, and in what ways, should we invest in attracting,
  onboarding, and working with contributors?}
\end{adjustwidth}

Setting contributor expectations accurately is one of the best moves a
project can make.  For example, an open source project is not
obligated to be Wide Open (\fullref{archetype:wide-open}), but it
should avoid accidentally misrepresenting itself as Wide Open if it's
not going to provide the kind of contributor support that a Wide Open
project would normally provide.  Potential contributors want to see
what they're getting into before they invest significant effort.  The
more clearly a project provides such information up front, the easier
it is for suitable contributors to find their way in and, equally
importantly, for unsuitable ones to self-select out.

Here are some of the questions developers typically have regarding
participation in an open source project:

\begin{itemize}
  \item I'm considering contributing.  What's the typical onboarding
    path?  How far can a third party contributor go?
  \item I am a new contributor.  How much influence should I expect to
    have in this project, if I continue contributing at my current
    level?  What about if I start contributing more frequently or more
    substantially?
  \item I am part of the maintainance team.  How much time should I
    expect to spend mentoring new contributors?
\end{itemize}

This section provides some sample public-facing language for setting
contributor and maintainer expectations.  Please note that the right
language for your particular project might be a customized combination
that draws parts from different samples, just as your project itself
might reflect a few different archetypes.

Where each project puts such language will vary, but a typical place
might be in a section called ``\texttt{\textbf{Contribution
    Expectations}}'' in the project's \texttt{README.md} or
\texttt{CONTRIBUTING.md} file.  The sample language offered here
refers to archetypes in order to clarify the reasons behind the
expectations being described, but does not assume that the reader is
familiar with archetypes.

\subsection{Sample contribution language: Rocket Ship to Mars}\label{sample-readme:rstm}

\begin{verbatim}
## Contribution Expectations

 This project is currently focused on high-velocity development toward
 a very specific goal: <summarize goal in one sentence here>.

 We welcome patches with bug fixes, and we're happy to answer
 questions if you've already put in some effort to find the answer
 yourself.  Please note, however, that we're unlikely to consider new
 feature contributions unless there's a strong argument that the
 proposed feature is fully in line with our stated goals.  If you're
 not sure, please feel free to ask.

 Our technical scope and project governance may open up later, of
 course, especially after we release version 1.0.  For now, though, we
 would characterize this project as being an example of the "Rocket
 Ship to Mars" archetype (see TBD URL for details about RStM and other
 open source archetypes).
\end{verbatim}

\subsection{Sample contribution language: Business to Business (B2B)}\label{sample-readme:b2b}

\begin{verbatim}
## Contribution Expectations

 If you have experience in <usage domain>, you should be able to get a
 development version of this up and running within a day, following
 the instructions in INSTALL.md.

 We very much welcome contributions of bugfixes and enhancements.  If
 your contribution might involve significant code change, or isn't
 clearly on the project roadmap, please start a discussion in
 <development forum> first -- we want to make sure we're all on the
 same page before you or your organization start to invest significant
 time and effort in technical work.

 Most of the development momentum in this project comes from enterprise
 users in <usage domain>.  Generally, their participation tends to be
 long-term and to align with their organization's business objectives.
 A given organization typically has several contributors working in
 the project at any given time.  While it is not necessary that your
 participation conform to this pattern, or that your motivations be
 similar to theirs, we think it's useful for potential contributors to
 know the environment: if your organization also has a long-term
 interest in this code base, you'll probably fit right in.

 (See the "Business to Business (B2B)" archetype at TBD URL for more
 about how this project compares to other open source projects.)
\end{verbatim}

\subsection{Sample contribution language: Multi-Vendor Infrastructure}\label{sample-readme:mvi}

\begin{verbatim}
## Contribution Expectations
  TBD
\end{verbatim}

\subsection{Sample contribution language: Wide Open}\label{sample-readme:wo}

\begin{verbatim}
## Contribution Expectations
  TBD
\end{verbatim}

\subsection{Sample contribution language: Mass Market}\label{sample-readme:mm}

\begin{verbatim}
## Contribution Expectations
  TBD
\end{verbatim}

\emph{TBD: For Controlled Ecosystem, Specialty Library, Trusted
  Vendor, and Upstream Dependency archetypes, way may handle them in
  various combinations, or in some other way discuss them together.
  With those, it's not necessarily the case that naming the archetype
  will be useful to contributors or maintainers, but anyway that will
  probably become clear as we try it.}
