\section{Setting Open Source Goals}\label{goals}

One of the keys to using open source as a tool is setting appropriate
goals.  Open source is well suited to encouraging a set of behaviors
and dynamics that are useful for solving a consistent range of
problems. This section contains a list of open source goals we find in
projects.  We intend it to be all-encompassing for typical use.  That
is, this list should capture the universe of goals open source
projects can reasonably achieve in the vast majority of circumstances.
If you have a goal that isn't captured by this list, it might be a
poor match for open source as it is typically practiced.\footnote{Is
  it also possible, of course, that the list is incomplete.  Please
  let us know if we missed something at \otsurl{https://github.com/OpenTechStrategies/open-source-archetypes/issues}.}

\textbf{Development and Collaboration}
\begin{itemize}
\item Expand or amplify developer base
\item Gain market insight
\item Gain or maintain insight in a particular technical domain
\item Influence a technical domain
\item Create a framework for partner collaboration
\item Lead a standardization effort
\item Disrupt an incumbent, hold off insurgents
  \end{itemize}
  
\textbf{External Positioning}
\begin{itemize}
\item Ease customer fears of vendor lock-in
\item Deepen engagement with users, create more paths for engagement
\item Transparency for customers and partners
\item Establish a basis for product reputation
\item Organizational branding and credibility
\item Product branding and credibility
  \end{itemize}

\textbf{Internal or Structural Change}
\begin{itemize}
\item Improve internal collaboration (cross-individual or cross-departmental)
\item Create opportunities for internal mobility
\item Expand or reshape developer hiring pool
\item Improve morale and retention
\item Create flow-paths for bottom-up innovation
\item Improve and maintain open source capabilities (technical and social)
  \end{itemize}

Before you start picking goals off this list, take a moment to reflect
on the overall goals of your effort.  Open source is one set of tools
in an overall strategy, and your open source goals only attain real
value in service to a larger vision.  Know your mission, your
cooperative and competitive landscape, your short and long-term goals,
and the resources you have at your command.  That context should tell
you a lot about what you need from your open source investment, and it
should help narrow your goals to an achievable list.

Start by reading the list over.  For the sake of convenience we have
divided the sixteen goals into three sections. If it is more useful
with a different grouping, do not hesitate to apply your own.  The
list ranges over a wide area.  It reflects a diverse set of practices
that can be fine-tuned in many directions.  It cannot, however, move
in sixteen directions at once.

You can use this list in a few ways.  First, read it over.  Think
about which of the goals would fill a gap in your larger strategy.

Second, match those goals up with archetypes that suit them.  If the
goals you've chosen align with archetypes that fit your overall
mission, that's a good sign you've found a path that has served other
projects successfully.  If you cannot find that alignment, think about
other goals, blended archetypes, or what context shifts would be
needed to enable your open source approach.

Finally, document your goals, from the large vision down to the open
source components and how those interact.  Make them explicit, and
share them with the team.  The more clarity people have about how the
work drives the strategy, the better they can fine tune the work to
your real goals.  Consider what indicators you can use to establish
starting points and targets.  We cover metrics in other material, but
you don't need quantifiable indicators at this stage.  You just need
to be able to answer the questions ``How will we know if it's
working?'' and ``What are the signs we need to make adjustments?''


