\hypertarget{preface}{}
\unnumberedsection{Preface}\label{preface}

Version 1 of this report was originally commissioned by the Mozilla
Corporation for internal purposes.  Mozilla wanted a shared framework
for discussing how to choose its open source investments projects.
For example, should a project prioritize development momentum over
early collaborator acquisition, or \foreignphrase{vice versa}?  Should
it optimize for individual contributors or for institutional partners?
Different projects will have different answers to these and many other
questions.

Realizing the benefit of having a common, public vocabulary of
archetypes, Mozilla decided to release the report publicly.  Very
little was changed between the internal version 1 and the public
version 1.  We left the Mozilla-specific orientation intact, on the
theory that the analysis would be clearer if tuned to a specific (and
well-known) organization rather than rewritten for a hypothetical
general audience.

The first edition stimulated a fair amount of discussion, both inside
and outside Mozilla.  In particular, in the course of applying
archetypes-based analysis to various open source projects, Mozilla
found places where the report could go into more depth or cover
certain adjacent topics.  Mozilla's thus commissioned an updated
Version 2.  The major differences are listed in the section
\fullref{v1-v2-changes} on p. \pageref{v1-v2-changes}.

As with Version 1, we believe that this is and should always be a work
in progress.  There are probably archetypes we have not thought of,
and for the archetypes we have identified there may be important
properties not discussed here.  When we published version 1 we wanted
it to be part of a larger conversation about open source project
archetypes, and fortunately that's what happened.  We intend the same
for Version 2, with the added desire that that conversation will grow
to include discussion of how projects can use archetype analysis to
guide their open source investments and overall strategy.
