\section{Transitions: Changing From One Archetype To Another}\label{transitions}

Projects can change from one archetype to another, and often do.  It
may be a deliberate and guided transformation, or it may be a gradual
and incremental evolution that is not driven by any one participant.

An example of a deliberate transition is Firefox, which went from
Rocket Ship To Mars (or something close to that) during its initial
push to a 1.0 release to a mode much closer today to Mass Market.

An example of a less centrally-driven and more incremental transition
is the WordPress open source project, which has been moving gradually
from something like Trusted Vendor to something more like Controlled
Ecosystem over a long period of time.  WordPress's transition has
occurred with the acquiescence and probably the encouragement of
WordPress's primary commercial backer, Automattic, Inc.  This is
probably the common case: with the exception of genuine hard forks, it
is rare for an archetype transition to occur entirely against the
wishes of the most-invested community members, although that does not
mean that every such transition is consciously driven by those members
either.

In general, transformations from Rocket Ship To Mars to some other
archetype are among the most common, because they reflect a natural
progression: from a small, focused, and unified core team working on
something new, to a position where more people are involved and have
broadening influence in the project.  If those new arrivals
are mostly developers whose motives are immediate rather than
strategic, that transformation might be toward Specialty Library.  If
they are developers using the code as part of other programs, it might
be toward Trusted Vendor or Upstream Dependency.  If they are plugin
developers and API consumers, it might be toward Controlled Ecosystem.
If they are organizations or organizationally-motivated developers,
B2B or Multi-Vendor Infrastructure might be likely places to end up.
If they are users, or a combination of users and developers, that
might point the way toward Mass Market or Wide Open.

Rocket Ship To Mars is not the only starting point for archetype
transitions.  A project might move from B2B to Upstream Dependency as
it becomes commoditized, for example, perhaps through competition or
perhaps just through rapid adoption.  We will not attempt here to map
out the full combinatoric range of possible archetype transitions.
Having such a map would not be very useful anyway.  What is useful
instead is to be always alert to the possibility of transformation,
and to ask, at strategic points in a project's life cycle, whether a
purposeful transformation might be called for, and whether there might
be bottom-up pressure pushing the project toward a particular new
archetype.
