\newpage
\section{Worksheet: Setting Goals and Fish}\label{worksheets}

\raggedright
\thispagestyle{empty}
\begin{center}
  \LARGE{Open Source Goal-Setting Worksheet}\\
  \vspace{1em}
  \normalsize{Version 1.0}
\end{center}
\vspace{2em}

\begin{multicols*}{2}

  \large{\textbf{Overall Goals}} \\
  \vspace{1em}
  
  In this column, describe your overall goals for your investment in
  this product or technology.
  
  \vspace{25em}
  
  \large{\textbf{Project Goals}}\\
  \vspace{1em}
  In this column, describe the goals of the open source project (not
  just your portion of the project).
  \vspace{22em}
\end{multicols*}

\newpage
\thispagestyle{empty}


\begin{textblock*}{3.5in}(0.5in,0.5in) % {block width} (coords)
  \large{\textbf{Your Open Source Goals}}\\
  \vspace{1em}

  On this page, circle or highlight up to three important open source
  goals from the list.  Select up to 3 more secondary goals.  Note
  them with a checkmark.
\end{textblock*}

\begin{textblock*}{3.5in}(3.75in,2.5in) % {block width} (coords)
\textbf{Development And Collaboration Goals} \\
Amplify or expand developer base\\
Market and contextual insight\\
Framework for partner collaboration\\
Lead a standardization effort\\
Disrupt an incumbent, hold off insurgents
\vspace*{2em}
\end{textblock*}

\begin{textblock*}{3.5in}(3.75in,4.65in) % {block width} (coords)
\textbf{External Marketing Goals}
Ease vendor lock-in fear \\
Engage with users \\
Transparency for customers and partners\\
Establish a basis for product reputation\\
Branding and credibility
\vspace*{2em}
\end{textblock*}

\begin{textblock*}{3.5in}(3.75in,6.5in) % {block width} (coords)
\textbf{Internal Goals} \\
Improve internal collaboration \\
Improve developer hiring pool \\
Improve morale and retention \\
Innovation \\
Improve open source capabilities
\end{textblock*}
\vspace*{2em}

\newpage
\thispagestyle{empty}
\begin{center}
  \LARGE{Open Source Ecosystem Mapping Worksheet}\\
  \vspace{1em}
  \normalsize{Version 1.0}
\end{center}
\vspace{2em}

Use this page to draw a map or directed graph of current and potential
actors in your ecosystem. List service providers and group them by the
type of service they offer.  Identify potential collaborators, and
mark the ones with competitive service offerings.  Identify competing
open source and proprietary substitutes for your open source project.
  \vspace{1em}

Place actors with large, current impact closer to the center of the
map and future recruits further away.  The open source project belongs
at the center, and your company might be close in or further out,
depending on its current effective scale of investment.
  \vspace{1em}

When done, note interesting relationships between various nodes on the
map.  Add customers in another color.  This map is a picture of your
world as it currently exists and how it might change in the near-term
future.  Be sure to save a snapshot of this map and see how it shifts
over time.

% https://twitter.com/dlshadothman/status/1065317596440989699
% is an example of an ecosystem map, but I'm not sure it's
% the right example for our purposes:
% 
% 1) It's based on publicly available data, whereas a real ecosystem
%    map would probably include privately-known information or
%    information that one is willing to guess about in private that
%    one wouldn't be as comfortable guessing about in public.  
%
% 2) It doesn't name very many names.  There are a few, but mostly it
%    talks in the abstract about communities, governments, etc.  That
%    can be useful for archetypes analysis too, but I think we want an
%    example that demonstrates the naming of specific actors.
%
% 3) It's done as a large, complex Venn diagram, which is a bit
%    limited in comparison to the ecosystem maps we usually use.
